---
title: "report"
author: "Stephen Cristiano"
date: "September 12, 2015"
output: html_document
---
Code for obtaining australian homicide data from the journal article is below, but will not be run in document as is behind a paywall and may not be accessible off the JHU network.

```{r fetch_data, echo=FALSE, eval=FALSE}
library(rvest)
library(plyr)
library(dplyr)
url = "http://injuryprevention.bmj.com/content/12/6/365/T2.expansion.html"
x = html(url)
tab = html_table(x)[[1]]
tab$Year = as.numeric(tab$Year)
tab = tab[-1, ]
tab = tab[-nrow(tab), ]

# using dplyr syntax
tab = filter(tab, !is.na(Year))

# renaming a column
tab = plyr::rename(tab, c("Year" = "year", "Population" = "population", 
                          "Firearm suicide" = "gun_suicide", "Firearm homicide" = "firearm_homicide",
                          "Homicide minus mass shootings" = "homicide_no_mass","Unintentional" = "unintentional",
                          "Undetermined" = "undetermined", "Total"= "total", "Total non-firearm homicide" = "all_non_gun_homicide",
                          "Suicide by all methods including firearm" = "all_suicide",
                          "Homicide by all methods including firearm" = "all_homicide",
                          "Total non-firearm suicide" = "other_suicide"))
tab$population <- gsub(" ", "", tab$population)
tab$gun_suicide = trimws(tab$gun_suicide)

# Splitting the string on spaces, then taking first element
# then turning that into a number
ss = strsplit(tab$firearm, " ")
ss = sapply(ss, function(x) x[1])

?ss = as.numeric(ss)

# Using a regular expression (regex) to extract the element we want
pop = gsub("(.*) \\(.*\\)", "\\1", tab$firearm)
df <- sapply(1:ncol(tab), function(x) as.numeric(gsub("(.*) \\(.*\\)", "\\1", tab[,x])))
colnames(df) <- colnames(tab)
df <- as.data.frame(df)
write.csv(df, "australia_longitudinal.csv", row.names=FALSE)
```

Reading in the data previously saved from the above chunk:

```{r read_data, echo=FALSE}
df <- read.csv("australia_longitudinal.csv")

### add treatment indicator (year of intervention)
df$treatment <- ifelse(df$year < 1997, 0, 1)

```
Next we plot homicide trends:
```{r plot_trends}
## create variables for homicide rate per 100k
gunhom <- df$homicide_no_mass/df$population * 1e5
allhom <- df$all_homicide/df$population * 1e5
otherhom <- df$all_non_gun_homicide/df$population * 1e5

## plot trends all on single graph
plot(df$year, allhom, type="l",ylim=c(0,2.4), main="Homicide trends", xlab="Year", ylab="Homicides per 100k")
lines(df$year, gunhom, lty=6, col="gray40")
lines(df$year, otherhom, lty=2, col="gray40")
legend("topright", c("All Homicides", "Firearm homicides", "Non-firearm homicides"), cex=0.5, lty=c(1,2,6), col=c("black", "gray40", "gray40"), bty="n")
abline(v=1997, col="tomato2")
```
Modeling code (note: requires package 'nlme' and 'lmtest'):
```{r segmented_regression, echo=FALSE}
## need nlme package and lmtest package
library(nlme)
library(lmtest)
### segmented regression
df$gunhomratio <- gunhom
df$allhomratio <- allhom
df$otherhom <- otherhom
df$time <- 1:nrow(df)
df$time2 <- df$time - 18
df$postslope <- ifelse(df$treatment==0, 0, df$time2)

## examine autocorrelation
acf(df$gunhomratio)
acf(df$allhom)
acf(df$otherhom)
```

```{r model_only_gun}
## linear model without 
## model <- lm(log(gunhomratio) ~ time + treatment + time:treatment, ## data=df)
## model

### we use the gls function in the nlme package to fit a linear model with autocorrelated errors 

## model allowing for differences in the slope
model2 = gls(log(gunhomratio) ~ time + treatment + postslope, data = df, correlation=corARMA(p=1))
model2
summary(model2)

## model with parallel slope pre and post intervention
model3 = gls(log(gunhomratio) ~ time + treatment, data = df, correlation=corARMA(p=1))
model3
summary(model3)

lrtest(model3, model2)
```
Doing analysis for overall homicide rate
```{r model_all}
### All homicide data
## model allowing for differences in the slope
model2.all = gls(log(allhom) ~ time + treatment + postslope, data = df, correlation=corARMA(p=1))
model2.all
summary(model2.all)

## model with parallel slope pre and post intervention
model3.all = gls(log(allhom) ~ time + treatment, data = df, correlation=corARMA(p=1))
model3.all
summary(model3.all)

lrtest(model3.all, model2.all)
##
### plot predicted trends pre and post intervention
pre.inv.all <- predict(model2.all, data.frame(treatment= 0, time = 1:18, postslope = 0))
post.inv.all <- predict(model2.all, data.frame(treatment = 1, time = 19:25, postslope = 1:7))



```
Making final plot of predicted outcomes:
```{r plotting}
##
### plot predicted trends pre and post intervention
pre.inv <- predict(model3, data.frame(treatment= 0, time = 1:18, postslope = 0))
post.inv <- predict(model3, data.frame(treatment = 1, time = 19:25, postslope = 1:7))

plot(log(gunhomratio) ~ time, data=df, cex=0.7, main="Predicted gun homicides", ylab="log(firearm homicide ratio)", type = "l", lwd=2, col="gray40")
lines(1:18,pre.inv, col = 'red', lwd = 2)
lines(19:25, post.inv, col = 'red', lwd = 2)

## extrapolate from pre intervention
pre.extrap <- predict(model3, data.frame(treatment= 0, time = 19:25, postslope = 1:7))
lines(19:25, pre.extrap, col = 'blue', lwd = 2, lty = 2)

### total homicide
pre.inv.all <- predict(model3.all, data.frame(treatment= 0, time = 1:18, postslope = 0))
post.inv.all <- predict(model3.all, data.frame(treatment = 1, time = 19:25, postslope = 1:7))

plot(log(allhom) ~ time, data=df, cex=0.7, type="l", lwd=2, col="gray40", main = "Predicted total homicides", xlab = "log(total homicide ratio")
lines(1:18,pre.inv.all, col = 'red', lwd = 2)
lines(19:25, post.inv.all, col = 'red', lwd = 2)

## extrapolate from pre intervention
pre.extrap.all <- predict(model3.all, data.frame(treatment= 0, time = 19:25, postslope = 1:7))
lines(19:25, pre.extrap.all, col = 'blue', lwd = 2, lty = 2)
```